---
title: "Code for the calculation of bipartite networks metrics in the ectoparasite  community of Planiliza haematocheilus for the Sea of Azov and Sea of Japan"
author: "Mar Llaberia-Robledillo, Juan A. Balbuena, Volodimir Sarabeev, Cristina Llopis-Belenguer"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 15, 
                      fig.align = "center")

```

# Libraries
Load the libraries required.
```{r}
library(bipartite)
library(parallel)

```
# Data
## Set a working directory
Optional but very useful step when you are working with multiple data files. It allows you to call files without writing their path, in addition to allowing data to be stored in the same folder, the working directory.
```{r}
setwd("~/Desktop/Planiliza haematocheilus (Lh)/ecto")
```

## Host-parasite abundance data
The main data necessary for the analysis belongs to a database of helminth parasites of the grey mullets (see main text for description). First, we edit the matrices separating those P. haematocheilus individuals belonging to the Sea of Azov - Black Sea and those belonging to the Sea of Japan. Second, we eliminate those data that are not necessary for our study (date, season, sex, etc.). Third, the data are divided according to the transmission strategy of the parasite species: ectoparasites, actively transmitted parasites, trophically transmitted parasites and finally, the whole community. 
This code shows the ectoparasite community analysis. We upload the host-parasite abundance matrices once arranged.
```{r}
azov_ecto<- read.table("azov_ecto.txt", header = T, row.names = 1)
japan_ecto<- read.table("japan_ecto.txt", header = T, row.names = 1)

```

## Real values of the Azov and Japan networks
We calculated three indices: connectance (C)(Dunne et al. 2002), weighted nesting based on overlap and decreasing fill (WNODF) (Almeida-Neto and Ulrich 2011), and modularity (Q) (Newman and Girvan 2004). The first two are calculated with a function (networklevel), while the third with another function (computeModules).In addition, we add a time calculator to determine the computation time needed later.
```{r}

## SEA OF AZOV 
t <- proc.time()
network_ecto_azov <- networklevel(azov_ecto, index = c("connectance", "weighted NODF"))
modularity_ecto_azov <- computeModules(azov_ecto)@likelihood
proc.time()-t

# join the values of the indices in the same object
index_ecto_azov <- data.frame((t(network_ecto_azov)), modularity_ecto_azov)
# (optional) save the results
# write.csv2(index_ecto_azov,file = "realvalues_azov.csv", row.names = FALSE)

## SEA OF JAPAN
t <- proc.time()
network_ecto_japan <- networklevel(japan_ecto, index = c("connectance", "weighted NODF"))
modularity_ecto_japan <- computeModules(japan_ecto)@likelihood
proc.time()-t

# join the values of the indices in the same object
index_ecto_japan <- data.frame((t(data.frame(network_ecto_japan))), modularity_ecto_japan)
# (optional) save the results
# write.csv2(index_ecto_japan,file = "realvalues_japan.csv")
```

### Bootstrap values
We generate 1,000 random networks from each matrix with replacement, row-based bootstrap replicas, and a fixed number of individuals. Then, we calculate the values of the indices for each of the replicated networks.
```{r}

## SEA OF AZOV
t <- proc.time()
# bootstrap funciton
bootSxSP <- function (azov_ecto) {
  a <- sample(seq_len(nrow(azov_ecto)), nrow(azov_ecto), replace = TRUE)
  azov_ecto <- azov_ecto[a, ]
  return(azov_ecto)
}
A_R <- replicate(1000, bootSxSP(azov_ecto), simplify = FALSE)

# calculation of index values for bootstrap results
cores <- detectCores()
cl <- makeForkCluster(20)
netlevelazov <- parSapply(cl, A_R, function(x) networklevel (x, index = c("connectance", "weighted NODF")))
modularity_azov<- parSapply(cl, A_R, function(x)computeModules(x)@likelihood)
stopCluster(cl)
proc.time()-t

# join the values of the indices in the same object
boot_ecto_azov <- data.frame(t(netlevelazov), modularity_azov)
# save the results 
write.csv2(boot_ecto_azov, file = "bootnet_ecto_azov.csv", row.names = FALSE)


## SEA OF JAPAN
t <- proc.time()
# bootstrap funciton
bootSxSP <- function (japan_ecto) {
  a <- sample(seq_len(nrow(japan_ecto)), nrow(japan_ecto), replace = TRUE)
  japan_ecto <- japan_ecto[a, ]
  return(japan_ecto)
}
A_R <- replicate(1000, bootSxSP(japan_ecto), simplify = FALSE)

# calculation of index values for bootstrap results
cores <- detectCores()
cl <- makeForkCluster(20)
netleveljapan <- parSapply(cl, A_R, function(x) networklevel (x, index = c("connectance", "weighted NODF")))
modularity_japan<- parSapply(cl, A_R, function(x)computeModules(x)@likelihood)
stopCluster()
proc.time()-t

# join the values of the indices in the same object
boot_ecto_japan <- data.frame((t(data.frame(netleveljapan))), modularity_japan)
# (optional) save the results 
write.csv2(boot_ecto_japan, file = "bootnet_ecto_japan.csv", row.names = FALSE)
```

### Nullmodel values 
We generate 1,000 replicas of the network matrices with the null.model function, with the swap.web and vaznull algorithms. Then, we calculate the values of the indices for each of the generated networks.
```{r}

## SEA OF AZOV
t <- proc.time()
## with swap.web algorithm (for WNDOF and Q)
nullazov_ecto_swap <- nullmodel(azov_ecto, N=1000, method = 2) #method 2/"swap.web"
## with vaznull algorithm (for WNODF only)
nullazov_ecto_vaz <- nullmodel(azov_ecto, N=1000, method = 3) #method 3/"vaznull"

cores <- detectCores()
cl <- makeForkCluster(20)

null_netlevelazov_s <- parSapply(cl, nullazov_ecto_swap, function(x) networklevel (x, index = c("weighted NODF")))
null_modularazov_s <-parSapply(cl, nullazov_ecto_swap, function(x) computeModules(x)@likelihood)

null_netlevelazov_v <- parSapply(cl, nullazov_ecto_vaz, function(x) networklevel (x, index = c("weighted NODF")))
null_modularazov_s <-parSapply(cl, nullazov_ecto_vaz, function(x) computeModules(x)@likelihood)

stopCluster(cl)
proc.time()-t

# join the values of the indices in the same object
null_swap_ecto_azov <- data.frame(data.frame(null_netlevelazov_s), null_modularazov_s)
null_vaz_ecto_azov <- data.frame(data.frame(null_netlevelazov_v), null_modularazov_v)

# (optional) save the results 
#write.csv2(null_swap_ecto_azov, file = "null_swap_ecto_azov.csv", row.names = FALSE)
#write.csv2(null_vaz_ecto_azov, file = "null_vaz_ecto_azov.csv", row.names = FALSE)


## SEA OF JAPAN
t <- proc.time()
## with swap.web algorithm (for WNDOF and modularity)
nulljapan_ecto_swap <- nullmodel(japan_ecto, N=1000, method = 2) #method 2/"swap.web"
## with vaznull algorithm (for WNODF only)
nulljapan_ecto_vaz <- nullmodel(japan_ecto, N=1000, method = 3) #method 3/"vaznull"

cores <- detectCores()
cl <- makeForkCluster(20)

null_netleveljapan_s <- parSapply(cl, nulljapan_ecto_swap, function(x) networklevel (x, index = c("weighted NODF")))
null_modularjapan_s <-parSapply(cl, nulljapan_ecto_swap, function(x) computeModules(x)@likelihood)

null_netleveljapan_v <- parSapply(cl, nulljapan_ecto_vaz, function(x) networklevel (x, index = c("weighted NODF")))
null_modularjapan_v <-parSapply(cl, nulljapan_ecto_vaz, function(x) computeModules(x)@likelihood)

stopCluster(cl)
proc.time()-t

# join the values of the indices in the same object
null_swap_ecto_japan <- data.frame(data.frame(null_netleveljapan_s), null_modularjapan_s)
null_vaz_ecto_japan <- data.frame(data.frame(null_netleveljapan_v), null_modularjapan_v)

# save the results 
#write.csv2(null_swap_ecto_japan, file = "null_swap_ecto_japan.csv", row.names = FALSE)
#write.csv2(null_vaz_ecto_japan, file = "null_vaz_ecto_japan.csv", row.names = FALSE)

```

## Standardize the real values of the network and the values of the null model
We standardized the real values of the networks and the values obtained from the 1,000 bootstrap networks of the WNODF and Q indices.
```{r}

# SEA OF AZOV
## WNODF values
index_ecto_azov [,2] #real value
boot_ecto_azov [,2] #bootstrap values 
# formula to standardize the networks values
stan_WNODF  <- function (x) {
  mean <- mean(null_swap_ecto_azov[,1])
  sd <- sd(null_swap_ecto_azov[,1])
  y <- (x)
  result <- (y-mean)/sd
}
a_w_value_wnodf <- sapply(index_ecto_azov[,2], stan_WNODF) #real standardized value 
boot_stanWNODF_azov <- as.data.frame(sapply(boot_ecto_azov[,3], stan_WNODF)) #bootstrap standardized values 
#write.csv2(boot_stanWNODF_azov, file = "boot_stanWNODF_azov.csv")
hist(boot_stanWNODF_azov)

## Q VALUES 
index_ecto_azov [,3] #real value
boot_ecto_azov [,3] #bootstrap values 
stan_Q  <- function (x) {
  mean <- mean(null_swap_ecto_azov[,2])
  sd <- sd(null_swap_ecto_azov[,2])
  y <- (x)
  result <- (y-mean)/sd
}
a_w_value_q <- sapply(index_ecto_azov[,3], stan_Q) #real standardized value 
boot_stanQ_azov <- as.data.frame(sapply(boot_ecto_azov[,7], stan_Q)) #bootstrap standardized values 
#write.csv2(boot_stanQ_azov, file = "boot_stanQ_azov.csv")
hist(boot_stanQ_azov)


# SEA OF JAPAN 
## WNODF VALUES
index_ecto_japan [,2] #real value
boot_ecto_japan [,2] #bootstrap values 
stan_WNODF  <- function (x) {
  mean <- mean(null_swap_ecto_japan[,1])
  sd <- sd(null_swap_ecto_japan[,1])
  y <- (x)
  result <- (y-mean)/sd
}
j_w_value_wnodf <- sapply(index_ecto_japan[,2], stan_WNODF) #real standardized value 
boot_stanWNODF_japan <- as.data.frame(sapply(boot_ecto_japan[,3], stan_WNODF)) #bootstrap standardized values 
#write.csv2(boot_stanWNODF_japan, file = "boot_stanWNODF_japan.csv")
hist(boot_stanWNODF_japan)

## Q VALUES 
index_ecto_japan [,3] #real value
boot_ecto_japan [,3] #bootstrap values
stan_Q  <- function (x) {
  mean <- mean(null_swap_ecto_japan[,2])
  sd <- sd(null_swap_ecto_japan[,2])
  y <- (x)
  result <- (y-mean)/sd
}
j_w_value_q <- sapply(index_ecto_japan[,3], stan_Q) #real standardized value 
boot_stanQ_japan <- as.data.frame(sapply(boot_ecto_japan[,7], stan_Q)) #bootstrap standardized values 
#write.csv2(boot_stanQ_japan, file = "boot_stanQ_japan.csv")
hist(boot_stanQ_japan)

```

## Statistical analysis: MANN - WHITNEY TEST
Statistical differences of the three indices between the native and invaded areas were tested by means of non-parametric Mann-Whitney-Wilcoxon tests.
```{r}

# C
wilcox.test(index_ecto_azov[,1], index_ecto_japan[,1], conf.int = TRUE)
# WNODF
wilcox.test(null_swap_ecto_azov[,1], null_swap_ecto_japan[,1], conf.int = TRUE)
# Q
wilcox.test(null_swap_ecto_azov[,2], null_swap_ecto_japan[,2], conf.int = TRUE)
wilcox.test((quantile(boot_stanq_azov, c(0.025,0.975))),(quantile(boot_stanq_japan, c(0.025,0.975))))


```

### Statistical metrics of the values of the indices of the Sea of Azov
We collected the real value, mean, standard deviation, and 95% confidence intervals of the indices for both areas.
```{r}

# SEA OF AZOV
## C VALUES
a_ecto_value_c <- index_ecto_azov[,1] #real value
a_ecto_mean_c <- mean(boot_ecto_azov[,1])
a_ecto_sd_c <- sd(boot_ecto_azov[,1])
a_ecto_ci_c <- quantile(boot_ecto_azov[,1], c(0.025,0.975))

## WNODF VALUES 
a_ecto_value_wnodf <- sapply(index_ecto_azov[,2], stan_WNODF) #real standardized value 
a_ecto_mean_wnodf <- mean(boot_stanWNODF_azov)
a_ecto_sd_wnodf <- sd(boot_stanWNODF_azov)
a_ecto_ci_wnodf <- quantile(boot_stanWNODF_azov, c(0.025,0.975))

## Q VALUES 
a_ecto_value_q <- sapply(index_ecto_azov[,3], stan_Q) #real standardized value
a_ecto_mean_q <- mean(boot_stanQ_azov)
a_ecto_sd_q <- sd(boot_stanQ_azov)
a_ecto_ci_q <- quantile(boot_stanQ_azov, c(0.025,0.975))


# SEA OF JAPAN 
## C VALUES 
j_ecto_value_c <- index_ecto_japan[,1] #real value
j_ecto_mean_c <- mean(boot_ecto_japan[,1])
j_ecto_sd_c <- sd(boot_ecto_japan[,1])
j_ecto_ci_c <- quantile(boot_ecto_japan[,1], c(0.025,0.975))

## WNODF VALUES 
j_ecto_value_wnodf <- sapply(index_ecto_japan[,2], stan_WNODF) #real standardized value 
j_ecto_mean_wnodf <- mean(boot_stanWNODF_japan)
j_ecto_sd_wndof <- sd(boot_stanWNODF_japan)
j_ecto_ci_wndof <- quantile(boot_stanWNODF_japan, c(0.025,0.975))

## Q VALUES 
j_ecto_value_q <- sapply(index_ecto_japan[,3], stan_Q) #real standardized value 
j_ecto_mean_q <- mean(boot_stanQ_japan)
j_ecto_sd_q <- sd(boot_stanQ_japan)
j_ecto_ci_q <- quantile(boot_stanQ_japan, c(0.025,0.975))


## Create a data frame with all the index results 

a_ecto_connectance <- c(a_ecto_value_c, a_ecto_mean_c, a_ecto_sd_c, a_ecto_ci_c)
j_ecto_connectance <- c(j_ecto_value_c, j_ecto_mean_c, j_ecto_sd_c, j_ecto_ci_c)

a_ecto_wnodf <-c(a_ecto_value_wnodf, a_ecto_mean_wnodf, a_ecto_sd_wnodf, a_ecto_ci_wnodf)
j_ecto_wnodf <-c(j_ecto_value_wnodf, j_ecto_mean_wnodf, j_ecto_sd_wnodf, j_ecto_ci_wnodf)

a_ecto_q <-c(a_ecto_value_q, a_ecto_mean_q, a_ecto_sd_q, a_ecto_ci_q)
j_ecto_q <- c(j_ecto_value_q, j_ecto_mean_q, j_ecto_sd_q, j_ecto_ci_q)

ecto_values <- rbind(a_ecto_connectance, j_ecto_connectance, a_ecto_wnodf, j_ecto_wnodf, a_ecto_q, j_ecto_q)
colnames(ecto_values) <- c("value", "mean", "sd", "ci")
ecto_values


```


